---
title: "Use CryptDiftR to replicate Biased drift paper"
author: "Ed Morrissey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicate_crypt_drift_paper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First load CryptDriftR.

```{r, echo=TRUE, message=FALSE}
library(CryptDriftR)
```


Next we load the data from Vermeulen, Morrissey et al Science 2013.

```{r}
data(PulseChaseData)
ls()
```
This data set comprises all 8 pulse chase data sets used in the paper. 

# Small intestine data analysis

## Wild type data
Fit the data with the neutral drift model.
```{r}
# Get time points from colnames
time_interval = data_SI_ahCreER_WT %>% colnames() %>% as.numeric()
# Fit model
fit_out_WT   = fitNeutralDrift(data_SI_ahCreER_WT, time_interval)
```
The model is fit using MCMC, so it's a good idea to check everything has gone well. This tends to not be a problem for this code, but worth checking.
```{r, fig.width = 7}
pp_convergence = plotsConvergence_Neutral(fit_out_WT)
plot(pp_convergence)
```

We can look at the parametres that were inferred.
```{r, fig.width = 8.5, fig.height= 7}
pp_posterior = plotPosterior_Neutral(fit_out_WT) 
plot(pp_posterior)
```

We can also access the optimal fit (MAP estimate is used for this).
```{r}
print(fit_out_WT$MAP_est)
```

It's also possible to plot the data with the optimal fit overlaid.

```{r, fig.width = 10, fig.height= 4}
plotsNeutralDrift_Fit(fit_out_WT)
```


## Mutant pulse chase data
Next we fit the biased drift model to the KRAS data.
```{r}
time_interval = data_SI_ahCreER_KRAS %>% colnames() %>% as.numeric()
# We use the WT parameters inferred from the WT data
fit_out_mut   = fitBiasedDrift(data_SI_ahCreER_KRAS, time_interval, WT_params = fit_out_WT$MAP_est)
```

Again we can check the convergence.
```{r, fig.width = 7}
pp_convergence =  plotsConvergence_Biased(fit_out_mut)
plot(pp_convergence)
```

We can plot the inferred parameter.

```{r, fig.width = 6}
pp_posterior = plotPosterior_Pr(fit_out_mut) 
plot(pp_posterior)
```

And calculate the estimate of the paramater

```{r}
print(fit_out_mut$Pr_quant)
```

We can also plot the fit

```{r , fig.width = 10, fig.height= 5}
pp_KRAS = plotsBiasDrift_Fit(fit_out_mut)
plot(pp_KRAS)
```

We can analyse the APC mutant data sets.
```{r, fig.width = 6, fig.height= 6}
time_interval = data_SI_ahCreER_APChet %>% colnames() %>% as.numeric()
fit_out_het   = fitBiasedDrift(data_SI_ahCreER_APChet, time_interval, WT_params = fit_out_WT$MAP_est)
time_interval = data_SI_ahCreER_APChom %>% colnames() %>% as.numeric()
fit_out_hom   = fitBiasedDrift(data_SI_ahCreER_APChom, time_interval, WT_params = fit_out_WT$MAP_est)
pp_het = plotPosterior_Pr(fit_out_het) + ggtitle("Apc-Het")
pp_hom = plotPosterior_Pr(fit_out_hom) + ggtitle("Apc-Hom")
grid.arrange(pp_het, pp_hom)
```


# Colon data analysis
First we get the neutral drift parameters for colon
```{r}
time_interval    = data_Colon_LGR5CreER_WT %>% colnames() %>% as.numeric() 
fit_out_colon_WT = fitNeutralDrift(data_Colon_LGR5CreER_WT, time_interval)
print(fit_out_colon_WT$MAP_est)
```

## Effect of p53
Next we analyse the effect of p53 mutation on colonic stem cells.

```{r, fig.width = 6}
time_interval   = data_Colon_LGR5CreER_p53 %>% colnames() %>% as.numeric()
fit_out   = fitBiasedDrift(data_Colon_LGR5CreER_p53, time_interval, WT_params = fit_out_colon_WT$MAP_est) 
pp_Pr_p53 = plotPosterior_Pr(fit_out) 
plot(pp_Pr_p53)
```


## Colitis
We calculate neutral drift parameters on a colitis background.
```{r, fig.width = 8.5, fig.height= 7}
time_interval   = data_DSSColon_LGR5CreER_WT %>% colnames() %>% as.numeric()
fit_out   = fitNeutralDrift(data_DSSColon_LGR5CreER_WT, time_interval) 
pp_post = plotPosterior_Neutral(fit_out)
plot(pp_post)
```

And next we estimate Pr for p53 mutant stem cells with a colitis background.

```{r, fig.width = 6}
time_interval = data_DSSColon_LGR5CreER_p53 %>% colnames() %>% as.numeric()
fit_out_mut   = fitBiasedDrift(data_DSSColon_LGR5CreER_p53, time_interval, WT_params = fit_out$MAP_est) 
pp_Pr_p53     = plotPosterior_Pr(fit_out_mut) 
plot(pp_Pr_p53)
```


